Vertical and oblique transmission with fluctuating transmission

### 2018-03-15 ms/fluctuating_oblique.tex
clean typos; added beta figure

In the case of $k=1$ we can also find the frequency of $A$ at the protected polymorphism, $x^*$.
Denote $F_B(F_A(x))-x = G(x) \cdot x (1-x) \cdot \gamma$ with
\begin{align} \label{eq:xstar_periodic_fluc_k=1}
G(x) = ax^2+bx+c, \\
\gamma = -\frac{1+xs}{s} \Big( 1 + s + s^2x (1-\rho) +s^2x^2(1-\rho) \Big) > 0, \\
a = s^2 \rho_1 (1-\rho_1) (1 - \rho_2 + \rho_2/\rho_1)) > 0, \\
b = s(1-\rho_1)(2\rho_2 - s\rho_1(1-\rho_2)), \\
c = \rho_2 - \rho_1 - s\rho_1(1-\rho_2),
\end{align} 
and then $x^*$ is a solution of the quadratic $G(x)=0$, given by $x^* = \frac{-b \pm \sqrt{b^2-4ac}}{2a}$.
The conditions for existence of the protected polymorphism \eqref{eq:poly_condition_periodic_fluc} is equivalent to $c<0$.% and $a+b+c>0$.
Therefore, $\sqrt{b^2-4ac} > b$, and if $0<s<1$ (such that $b$ is guaranteed to be positive) then $x^*= \frac{-b+\sqrt{b^2-4ac}}{2a}$ (\autoref{fig:rho1_rho2_k=1}).

@@

Second, if $s_t$ is symmetric around zero ($E[s_t]=E[s_t/(1+s_t)]=0$), and $\rho_t$ and $s_t$ are positively (negatively) correlated, such that vertical transmission is more likely when $A$ ($B$) is favored, then $E[-\rho_t s_t/(1+s_t)]<0$ ($E[-\rho_t s_t]<0$) and therefore fixation of $A$ ($B$) is stochastically locally stable.
For example, let $s_t \sim U(-1,1)$ and $\rho_t \sim \beta(1+s, 1)$, such that the covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 0.22$), that is, vertical transmission is more likely when $A$ if favored (when $s_t>0$) and oblique transmission is more  likely when $B$ is favored (when $s_t<0$), then ($Beta$ being the beta function), 
\begin{equation}
E[log(1+\rho_t s_t)] = \frac{1}{2} \int_{-1}^{1} {\int_0^1 {\frac{\rho^s \log{(1+\rho s)}}{Beta(1+s, 1)}  \; d\rho}\; ds} \approx 0.2525,
\end{equation}
and $B$ almost surely doesn't fix, and
\begin{equation}
E[log(1-\rho_t s_t / (1+s_t)] = \frac{1}{2} \int_{-1}^{1} {\int_0^1 {\frac{(1-\rho)^{-s/(1+s)} \log{(1-\rho s / (1+s))}}{Beta(1, 1-s/(1+s)}  \; d\rho}\; ds} \approx -0.076,
\end{equation}
so fixation of $A$ is stochastically locally stable.
The opposite occurs if $\rho_t \sim \beta(1+s, 1)$ and therefore the covariance of $s_t$ and $\rho_t$ is negative ($cov(s_t, \rho_t) \approx -0.2$); in that case, fixation of $B$ is stochastically locally stable and $A$ almost surely doesn't fix.

>>>>

In the case of $k=1$ we can also find the frequency of $A$ at the protected polymorphism, $x^*$.
Denote $F_B(F_A(x))-x = G(x) \cdot x (1-x) \cdot \gamma$ with
\begin{align} \label{eq:xstar_periodic_fluc_k=1}
G(x) = ax^2+bx+c, \\
\gamma = -\frac{1+xs}{s} \Big( 1 + s + s^2x (1-\rho) +s^2x^2(1-\rho) \Big) > 0, \\
a = s^2 \rho_1 (1-\rho_1) (1 - \rho_2 + \rho_2/\rho_1)) > 0, \\
b = s(1-\rho_1)(2\rho_2 - s\rho_1(1-\rho_2)), \\
c = \rho_2 - \rho_1 - s\rho_1(1-\rho_2),
\end{align} 
and then $x^*$ is a solution of the quadratic $G(x)=0$, given by $x^* = \frac{-b \pm \sqrt{b^2-4ac}}{2a}$.
The condition \eqref{eq:poly_condition_periodic_fluc} guarantees that $x^* \in (0,1)$ and is equivalent to $c<0$.% and $a+b+c>0$.
Therefore, $\sqrt{b^2-4ac} > b$, and if $0<s<1$ (such that $b$ is guaranteed to be positive) then $x^*= \frac{-b+\sqrt{b^2-4ac}}{2a}$ (\autoref{fig:rho1_rho2_k=1}).

@@

Second, 
%if $s_t$ is symmetric around zero ($E[s_t]=0$), and $\rho_t$ and $s_t$ are negatively correlated, such that vertical transmission is more likely when $B$ is favored, then $E[-\rho_t s_t]<0$ and therefore fixation of $B$ is stochastically locally stable.
let $s_t \sim U(-1,1)$ and $\rho_t \sim \beta(1+s, 1)$ (\autoref{fig:beta}), such that the covariance of $s_t$ and $\rho_t$ is positive ($cov(s_t, \rho_t) \approx 0.22$), that is, vertical transmission is more likely when $A$ if favored (when $s_t>0$) and oblique transmission is more  likely when $B$ is favored (when $s_t<0$), then ($Beta$ being the beta function), 
\begin{equation}
E[log(1+\rho_t s_t)] = \frac{1}{2} \int_{-1}^{1} {\int_0^1 {\frac{\rho^s \log{(1+\rho s)}}{Beta(1+s, 1)}  \; d\rho}\; ds} \approx 0.2525,
\end{equation}
and $B$ almost surely doesn't fix, and
\begin{equation}
E[log(1-\rho_t s_t / (1+s_t)] = \frac{1}{2} \int_{-1}^{1} {\int_0^1 {\frac{(1-\rho)^{-s/(1+s)} \log{(1-\rho s / (1+s))}}{Beta(1, 1-s/(1+s)}  \; d\rho}\; ds} \approx -0.076,
\end{equation}
so fixation of $A$ is stochastically locally stable.
The opposite occurs if $\rho_t \sim \beta(1+s, 1)$ and therefore the covariance of $s_t$ and $\rho_t$ is negative ($cov(s_t, \rho_t) \approx -0.2$); in that case, fixation of $B$ is stochastically locally stable and $A$ almost surely doesn't fix.

\begin{figure*}[hbt]
\centering
\includegraphics[width=\linewidth]{../figures/{beta}.pdf}
\caption{
\textbf{Beta distribution.}
\textbf{(A)} Distribution of $\rho_t \sim Beta(1+s_t, 1)$ for different values of $s_t$.
It can be seen that for $s_t>0$ vertical transmission is more likely, and for $s_t<0$ oblique transmission is more likely.
\textbf{(B)} The compound distribution of $\rho_t \sim Beta(1+s_t, 1)$ for $s_t \sim U(-1,1)$, demonstrating that, overall, oblique transmission is more likely than vertical transmission.
\textbf{(C)} The joint distribution of $\rho_t$ and $s_t$ demonstrates a positive correlation.
}
\label{fig:beta}
\end{figure*}
